---
title: "Solution for fitting session on BDA"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstanarm)
library(brms)
library(tidyverse)
```

## Case 1: fitting mixed effect models with rstanarm

### Code demo

```{r}
# load the dataset
dat <- read.csv("../data/fitting_01.csv")


# first model
m_1 <- stan_lmer(height ~ rain + (1 | grp), dat)

# summary
summary(m_1)

# check priors
prior_summary(m_1)

# check convergence
plot(m_1, plotfun = "trace", pars = c("(Intercept)", "rain", "sigma", "Sigma[grp:(Intercept),(Intercept)]"))
rhat(m_1)

# make pp_check
pp_check(m_1, nreps = 25)

# plot the predicted regression lines
newdat <- expand.grid(rain = seq(-2, 2, length.out = 10), grp = letters[1:n_grp])

pred <- posterior_predict(m_1, newdata = newdat)
## get median and 90 CrI
pred_sum <- apply(pred, 2, function(x) quantile(x, probs = c(0.05, 0.5, 0.95)))
newdat$height <- pred_sum[2,]
newdat$LCI <- pred_sum[1,]
newdat$UCI <- pred_sum[3,]

# plot
ggplot(newdat, aes(x=rain,y=height,ymin=LCI,ymax=UCI, group = grp)) +
  geom_ribbon(aes(fill=grp), alpha = 0.1) +
  geom_path(aes(color=grp)) 

# get R2 and loo
quantile(bayes_R2(m_1), probs = c(0.5))

ll_1 <- loo(m_1)

# a better plot
ggplot(newdat, aes(x=rain,y=height, group = grp)) +
  geom_ribbon(aes(fill=grp, ymin=LCI,ymax=UCI), alpha = 0.1) +
  geom_path(aes(color=grp)) +
  geom_point(data=dat, aes(color = grp))

```

### Exercices

Fit a random slope model to the data and extract:

- convergence checks
- posterior predictive checks
- plot of predicted regression lines with 80% Credible interval
- compare this model to the previous one using LOO and R2

```{r}

```

For the advanced people:

- remove the automatic scaling of the priors on the intercept, the slope and the deviations (sigma), compare the results to the other models
- compute the posterior probability that the effect of rain is larger for group a than for group b
- re-fit the model with uncorrelated random effects (check http://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#model-specification)


## Case 2: fitting varying variance model with brms

### code demo

```{r}
# load the data
dat_2 <- read.csv("../data/fitting_02.csv")

# fit the model
brm_form <- brmsformula(biomass ~ nutrient, sigma ~ nutrient)
m_3 <- brm(brm_form, data = dat_2)

# model summary
summary(m_3)

# check convergence
plot(m_3)
rhat(m_3)

# check posterior predictive distribution
brms::pp_check(m_3, type = "dens_overlay", nsamples = 25)

# plot 
mm <- marginal_effects(m_3, method = "predict")
plot(mm, points = TRUE)
```

### Exercices

- Identify the problem with the previous model and correct it in a new model
- Run the usual model checks (convergence, posterior predictive distribution ...)
- Plot the fitted regression line
- Compare the two models

```{r}

```

For the advanced people

- The previous plot show the prediction interval, make a plot with both the interval for the fitted slope and the prediction interval
- Change the slope for the nutrient effect to a cauchy distribution with mean 0 and deviation of 2
- Get the posterior probability that the variation at the both ends of the nutrient gradient are higher than in the middle of the gradient

## Case 3: fitting temporal autocorrelation with brms

```{r}
# read the data
dat_3 <- read.csv("../data/fitting_03.csv")
# check for autocorrelation
acf(dat_3$biomass)

# fit the model
m_5 <- brm(biomass ~ 1, data = dat_3, autocor = cor_ar(formula = ~ time, p = 1))

# model summary
summary(m_5)

# model checks
plot(m_5)
rhat(m_5)

# check posterior predictive distribution
brms::pp_check(m_5, type = "dens_overlay", nsamples = 25)

# plot fitted and predicted lines
newdat <- expand.grid(time = 1:n_step) # a new dataframe to project the model predictions

ff <- fitted(m_5, newdata = newdat) # the fitted (linear predictor) values from the model
pp <- predict(m_5, newdata = newdat) # the model prediction (fitted values + residual variations)
newdat <- cbind(newdat, pp, ff[,3:4])
names(newdat)[c(2,4:7)] <- c("biomass", "p_lci", "p_uci", "f_lci", "f_uci")

ggplot(newdat, aes(x = time, y = biomass)) +
  geom_ribbon(aes(ymin = p_lci, ymax = p_uci), alpha = 0.1) +
  geom_ribbon(aes(ymin = f_lci, ymax = f_uci), alpha = 0.1) +
  geom_path() + geom_path(data = dat_3)
```

### Exercices 

- Fit a model with a lag of 2
- Run the usual checks
- Compare with the previous model

```{r}

```

For the advanced people:

- Feierabend !